Structured Versus Decorated Cospans

### 2021-06-18 cospans_new.tex
Fixed more typos in the proof of Theorem 2.1

In more detail, if $(\phi,\phi_0)$ is the lax monoidal structure of the pseudofunctor $F$ as recalled in \cref{subsec:bicats}, the induced monoidal structure on the total Grothendieck category $\X=\inta F$ (\cref{def:GrothCat}) is given by
\begin{equation}\label{eq:explicitstructure2}
\Big(a,x\in F(a)\Big)\ot_{\X}\Big(b,y\in F(b)\Big)=\Big(a+b,\phi_{a,b}(x,y)\in F(a+b)\Big), \qquad I_{\X}=\Big(0_A,\phi_0\Big)
\end{equation}
If $F$ is a symmetric lax monoidal pseudofunctor, the induced monoidal structure in $\inta F$ is symmetric via
$$\left(\beta_{a,b},(u_{a,b})_{x,y}\right)\maps \left(a+b,\phi_{a,b}(x,y)\right)\simrightarrow\left(b+a,\phi_{b,a}(y,x)\right)$$ where $\beta$ is the canonical symmetry for $\A$ and $u$ is the natural isomorphism of \cref{eq:braidedpseudofun}.

@@

Now we are ready to use Shulman's result to prove  \cref{thm:decorated_cospans_1} and  \cref{thm:decorated_cospans_2}.  An $F$-decorated cospan from $a \in \A$ to $b \in \A$
is a cospan 
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B);
\end{tikzpicture}
\]
in $\A$ together with an object $x \in F(m)$.   This can be seen as a pair $(e \to m, x \in F(m))$ such that $e$ equals $a + b$.   But $(m,x \in F(m))$ is precisely an object of the Grothendieck category $\inta F$, and the opfibration $U \maps \inta F \to \A$ sends this object to $m$.  Thus, $(e \to m, x \in F(m))$ can be seen as an object of the comma category $\A/U$.   Let $\Phi \maps \A/U \to \A$ be the functor sending any object $(e \to m,x \in F(m)) \in \A/U$ to $e$.  Then an $F$-decorated cospan from $a$ to $b$ amounts to an object $c \in \A/U$ such that $\Phi(c) = a + b$.    With this choice of $\Phi$, Shulman's result gives the double category of $F$-decorated cospans.

@@

To apply \cref{lem:shulman} it suffices to prove that $\Phi$ is a Beck--Chevalley monoidal opfibration.   First, there is a monoidal structure on $\A/U$ coming from the monoidality of $U$.  On objects this is given by
\[  (a \xrightarrow{i} b, x) \otimes (a' \xrightarrow{i'} b', x') = (a+a' \xrightarrow{i+i'} b+b', 
\phi_{a,b}(d,d') )\]
where $\phi$ is the laxator for $F$.  The monoidal unit is $(0_\A \xrightarrow{!} 0_\A, x)$
where $x \in F(0_\A)$ is the element given by $\phi_0 \maps \one \to F(0_\A)$.   With this monoidal structure it is easy to see that $\Phi$ is strict monoidal.

@@

is the following map of $F'$-decorated cospans $\lH(\alpha)$ in $F'\lCsp$:

@@

where the decoration morphism $\tau$ is mapped to a decoration morphism as follows:

@@

\node (K) at (-1,-0.5) {$\mapsto$};

@@

\node (F) at (4,-1) {$F^\prime(H(n))$};

@@

(A) edge node [above] {$\phi$} (L)
(H) edge node [right] {$F(h)$} (I)

@@

\begin{thm}
\label{thm:functoriality}
Given two categories $\A$ and $\A'$ with finite colimits, two lax monoidal pseudofunctors $F, F' \maps (\A',+) \to (\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and a {\john monoidal???} natural isomorphism $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\Swarrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
the triple $(H,E,\theta)$ induces a double functor $\lH \maps F\lCsp \to F'\lCsp$ as defined above.
\end{thm}

@@

We described $\A/U$ and its monoidal structure in the proof of \cref{thm:decorated_cospans_1}, 
and of course the case of $\A'/U'$ is analogous.   We define the strong monoidal functor $G \maps \A/U \to \A'/U'$ as follows.  First we define $G$ on an object $(a \xrightarrow{i} b, x)$ of $\A/U$.
We can treat $x \in F(b)$ as a morphism $x \maps 1 \to F(b)$.    Then we define
\[  G(a \xrightarrow{i} b, x) := (H(a) \xrightarrow{H(i)} H(b), x') \]
where the object $x' \in F'(H(b))$ is picked out by the composite
\[  1 \xrightarrow{\phi_0} E(1) \xrightarrow{E(x)} E(F(b)) \xrightarrow{\;\theta_b\;} F'(H(b)) .\]
Then we define $G$ on a morphism
\[   (f,g,h) \maps (a \xrightarrow{i} b, x) \; \to \; (a' \xrightarrow{i'} b', x') \]
in $\A/U$.   For this we set
\[   G(f,g,h) = (H(f),H(g),???) \]

where $\tau$ and $\theta_h E(\tau) \phi$ are given respectively by:
\[
\begin{tikzpicture}[scale=1.5]
\node (G) at (-3.5,-0.5) {$1$};
\node (H) at (-2,0) {$F(m)$};
\node (I) at (-2,-1) {$F(n)$};
\node (J) at (-2.5,-0.5) {$\scriptstyle\Downarrow\tau$};
\node (A) at (-0.25,-0.5) {$1$};
\node (L) at (0.5,-0.5) {$E(1)$};
\node (B) at (2,0) {$E(F(m))$};
\node (D) at (2,-1) {$E(F(n))$};
\node (C) at (1.5,-0.5) {$\scriptstyle\Downarrow E(\tau)$};
\node (E) at (4,0) {$F^\prime(H(m))$};
\node (F) at (4,-1) {$F^\prime(H(n))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node [above] {$\phi$} (L)
(G) edge node [above] {$p$} (H)
(G) edge node [below] {$q$} (I)
(H) edge node [right] {$F(h)$} (I)
(B) edge node [above] {$\theta_m$} (E)
(D) edge node [above] {$\theta_n$} (F)
(E) edge node [right] {$F'(H(h))$} (F)
(L) edge node[above]{$E(p)$} (B)
(L) edge node[below]{$E(q)$} (D)
(B) edge node[right]{$E(F(h))$} (D);
\end{tikzpicture}
\]

@@

This theorem should generalize to the case when $\theta$ is a pseudonatural equivalence, but the weaker version suffices for our application in \cref{subsec:petrirates}.

>>>>

In more detail, if $(\phi,\phi_0)$ is the lax monoidal structure of the pseudofunctor $F$ as recalled in \cref{subsec:bicats}, the induced monoidal structure on the Grothendieck category $\X=\inta F$ (\cref{def:GrothCat}) is given by
\begin{equation}\label{eq:explicitstructure2}
\Big(a,x\in F(a)\Big)\ot_{\X}\Big(b,y\in F(b)\Big)=\Big(a+b,\phi_{a,b}(x,y)\in F(a+b)\Big), \qquad I_{\X}=\Big(0_A,\phi_0\Big)
\end{equation}
If $F$ is a symmetric lax monoidal pseudofunctor, the induced monoidal structure in $\inta F$ is symmetric via
$$\left(\beta_{a,b},(u_{a,b})_{x,y}\right)\maps \left(a+b,\phi_{a,b}(x,y)\right)\simrightarrow\left(b+a,\phi_{b,a}(y,x)\right)$$ where $\beta$ is the canonical symmetry for $\A$ and $u$ is the natural isomorphism of \cref{eq:braidedpseudofun}.

@@

Now we are ready to use Shulman's result to prove  \cref{thm:decorated_cospans_1} and  \cref{thm:decorated_cospans_2}.  An $F$-decorated cospan from $a$ to $b$ 
is a cospan 
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B);
\end{tikzpicture}
\]
in $\A$ together with an object $x \in F(m)$.   This can be seen as a pair $(e \to m, x \in F(m))$ such that $e$ equals $a + b$.   But $(m,x \in F(m))$ is precisely an object of the Grothendieck category $\inta F$, and the opfibration $U \maps \inta F \to \A$ sends this object to $m$.  Thus, $(e \to m, x \in F(m))$ can be seen as an object of the comma category $\A/U$.   Let $\Phi \maps \A/U \to \A$ be the functor sending any object $(e \to m,x \in F(m)) \in \A/U$ to $e$.  Then an $F$-decorated cospan from $a$ to $b$ amounts to an object $c \in \A/U$ such that $\Phi(c) = a + b$.    With this choice of $\Phi$, Shulman's result gives the double category of $F$-decorated cospans.

@@

To apply \cref{lem:shulman} it suffices to prove that $\Phi$ is a Beck--Chevalley monoidal opfibration.   First, there is a monoidal structure on $\A/U$ coming from the monoidality of $U$.  On objects this is given by
\[  (a \xrightarrow{i} b, x) \otimes (a' \xrightarrow{i'} b', x') = (a+a' \xrightarrow{i+i'} b+b', 
\phi_{a,b}(x,x') )\]
where $\phi$ is the laxator for $F$.  The monoidal unit is $(0_\A \xrightarrow{!} 0_\A, 0)$
where $0 \in F(0_\A)$ is the object given by $\phi_0 \maps \one \to F(0_\A)$.   With this monoidal structure it is easy to see that $\Phi$ is strict monoidal.

@@

is the following map of $F'$-decorated cospans in $F'\lCsp$:

@@

where the decoration morphism $\tau$ is mapped to a decoration morphism $\lH(\tau)$ as follows.  We can think of $\tau$ as a natural transformation

@@

(H) edge node [right] {$F(h)$} (I);
and in the same style of notation $\lH(\tau)$ is given by

@@

\node (F) at (4,-1) {$F^\prime(H(n))$.};

@@

(A) edge node [above] {$\phi_0$} (L)

@@

The square at right commutes strictly because $\theta$ is a natural isomorphism.   Perhaps we could generalize our construction to handle the case where $\theta$ is a pseudonatural equivalence, making this square commute up to a natural isomorphism.  However, the following result suffices 
for the application in \cref{subsec:petrirates}.

\begin{thm}
\label{thm:functoriality}
Given categories $\A$ and $\A'$ with finite colimits, lax monoidal pseudofunctors $F, F' \maps (\A',+) \to (\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and a {\john monoidal???} natural isomorphism $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\Swarrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
the triple $(H,E,\theta)$ induces a double functor $\lH \maps F\lCsp \to F'\lCsp$ defined as above.
\end{thm}

@@

We described $\A/U$ and its monoidal structure in the proof of \cref{thm:decorated_cospans_1}, 
and of course the case of $\A'/U'$ is analogous.   We define $G \maps \A/U \to \A'/U'$ as follows.  First we define $G$ on any object $(a \xrightarrow{i} b, x)$ of $\A/U$ as follows:
\[  G(a \xrightarrow{i} b, x) \; = \; (H(a) \xrightarrow{H(i)} H(b), \underline{x}). \]
To define $\underline{x}$, we treat the object $x \in F(b)$ as a functor 
$x \maps 1 \to F(b)$, and let $\underline{x} \in F'(H(b))$ be given by this functor:
\[  1 \xrightarrow{\phi_0} E(1) \xrightarrow{E(x)} E(F(b)) \xrightarrow{\;\theta_b\;} F'(H(b)) .\]
Next we define $G$ on any morphism
\[   (f,g,h) \maps (a \xrightarrow{i} b, x) \; \to \; (a' \xrightarrow{i'} b', x') \]
in $\A/U$.   For this we set
\[   G(f,g,h) = (H(f),H(g),\underline{h}) .\]
To define $\underline{h}$, we treat $h \maps F(g)(x) \to x'$ as a natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (G) at (-3.5,-0.5) {$1$};
\node (H) at (-2,0) {$F(b)$};
\node (I) at (-2,-1) {$F(b')$};
\node (J) at (-2.5,-0.5) {$\scriptstyle\Downarrow\tau$};
\path[->,font=\scriptsize,>=angle 90]
(G) edge node [above] {$x$} (H)
(G) edge node [below] {$x'$} (I)
(H) edge node [right] {$F(g)$} (I);
\end{tikzpicture}
\]
and let $\underline{h} \maps F'(H(g))(\underline{x}) \to \underline{x}'$ be given by this
natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (-0.25,-0.5) {$1$};
\node (L) at (0.5,-0.5) {$E(1)$};
\node (B) at (2,0) {$E(F(b))$};
\node (D) at (2,-1) {$E(F(b'))$};
\node (C) at (1.5,-0.5) {$\scriptstyle\Downarrow E(\tau)$};
\node (E) at (4,0) {$F^\prime(H(b))$};
\node (F) at (4,-1) {$F^\prime(H(b'))$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node [above] {$\phi_0$} (L)
(B) edge node [above] {$\theta_b$} (E)
(D) edge node [above] {$\theta_{b'}$} (F)
(E) edge node [right] {$F'(H(h))$} (F)
(L) edge node[above]{$E(x)$} (B)
(L) edge node[below]{$E(x')$} (D)
(B) edge node[right]{$E(F(h))$} (D);
\end{tikzpicture}
\]
One can check that with these definitions $G$ is a functor.

To make $G$ into a strong monoidal functor, we need to equip it with a natural isomorphism

@@
